The Shape Matching Element Method: Direct Animation of Curved Surface Models

### 2021-01-27 paper/sections/methods.tex
updated algorithm figure

\ty{Update notation, split this into two figures}
\begin{algorithm}[h]
    \label{alg:sem}
	\caption{Shape Matching Element Algorithm}\label{sem_algorithm}
	\begin{algorithmic}[1]
		\Procedure{Preprocess}{$model$}
		\State{// Get initial DOF and sample boundary}
		\State $\mathbf{q^0} \gets \text{GetControlPoints}(model)$ \Comment{Section 4}
		\State $\mathbf{J}, \mathbf{W} \gets \text{SampleNURBS}(model)$ \Comment{Section 4}
		\State $\mathbf{x^0} \gets \mathbf{Jq^0}$
		\State{$\mathbf{q} \gets \mathbf{q^0}$}
      \Statex

    	\State{// Generate material points inside the model}
		\State $\mathbf{X} \gets \text{RaycastQuadrature}(model)$ \Comment{Section 4.5}
		\Statex

      \State{// Weights for quadrature and boundary points}
     	\State $\mathbf{w} \gets \text{BlendingWeights}(model, \mathbf{X}, d_{\text{max}})$    \Comment{Section 4.3}
      	\State $\mathbf{w^*} \gets \text{BlendingWeights}(model, \mathbf{x^0}, d_{\text{max}})$   
		\Statex

		\State{// Generate centers of mass}
		\State $\bar{\mathbf{X}} \gets \text{GenerateCOM}(\mathbf{X}, \mathbf{w})$  \Comment{Section 4.}
		\Statex

		\State{// Create projection operator}
		\State $\Pi \gets \text{ShapeMatching}(\mathbf{q^0},\mathbf{J}, \mathbf{W},\bar{\mathbf{X}})$     	\Comment{Section 4.2}
		\Statex

		\State{// Form weighted monomial basis and mass matrices}
		\State $\mathbf{Y} \gets \text{WeightedMonomials}(\mathbf{X}, \mathbf{\bar{X}}, \mathbf{w})$   \Comment{Eqn 32}     	
		\State $\mathbf{Y^*} \gets \text{WeightedMonomials}(\mathbf{x^*}, \mathbf{\bar{X}}, \mathbf{w})$
    	\State $\mathbf{M} \gets \text{MassMatrix}(\mathbf{Y}, \Pi)$			\Comment{Section 4.6}
     	\State $\mathbf{M_E} \gets \text{ErrorMatrix}(\mathbf{Y^*}, \Pi)$	\Comment{Section 4.7}
     	\Statex
     	\myworries{todo: add volume}
     	\While {$simulating$}
     		\State{$\Pcox \gets \Pi\mathbf{q}$}
     		\ForAll{$i \gets 1$ to $|\mathbf{X}|$}
     			\State{$\mathbf{F}_i \gets \frac{\partial \mathbf{F}}{\partial c} c$}
				\State{$\frac{\partial V}{\partial c} \gets \frac{\partial V}{\partial c} +
				\frac{\partial \mathbf{F}}{\partial c}^T \frac{\partial \psi}{\partial \mathbf{F}}(\mathbf{F}_i)$}

				\State{$\frac{\partial^2 V}{\partial c^2} \gets \frac{\partial^2 V}{\partial c^2} +
				 \frac{\partial \mathbf{F}}{\partial c}^T 
				 \frac{\partial^2 \psi}{\partial \mathbf{F}^2}(\mathbf{F}_i)
				 \frac{\partial \mathbf{F}}{\partial c}
				 $}
  			\EndFor
  			\State{$\mathbf{K} \gets \Pi^T \frac{\partial^2 V}{\partial c^2} \Pi$}
  			\State{$f_{\text{internal}} \gets \Pi^T \frac{\partial V}{\partial c}$}
  			\State{$f_{\text{error}} \gets \mathbf{M_E}\mathbf{q}$}
  			\State{$f_{\text{total}} \gets f_{\text{internal}} + f_{\text{error}} + f_{\text{ext}}$}
  			\State{$\dot{\mathbf{q}} \gets \text{TimeIntegration}
  			(f_{\text{total}},\mathbf{K}, \dot{\mathbf{q}}, \mathbf{J} , \mathbf{M}, \mathbf{M_E})$}
  			\State{$\mathbf{q} \gets \mathbf{q} + h\dot{\mathbf{q}}$}
     	\EndWhile
	\EndProcedure
\end{algorithmic}
\caption{Shape Matching Element Algorithm}
\end{algorithm}

%\section{Methods}
\section{Geometric Model}

@@

\section{Physics-Based Animation}

Having defined the individual components of the SEM method, we now briefly detail how we apply it to the problem of 
physics-based animation, specifically elastodynamics. 
The advantage of having a unified kinematic description~(\refeq{kinematics}) is that we can directly apply Lagrangian Mechanics~\dave{cite Lanczos again} to arrive at 
the pertinent equations of motion:
\begin{equation*}
    \MM\ddot{\vc{q}} = \fo\left(\vc{q}\right) + \fext,
\end{equation*} where $\MM\in\real^{n\times n}$ is the mass matrix, $\vc{q}\in\real^{n}$ is the stacked vector of surface control points, $\fo\in\real^{n}$ is the 
vector of internal forces and $\fext\in\real^{n}$ are external forces such due to gravity or contact. 
We use penalty springs~\cite{10.1145/1964921.1964932} to resolve contact, and the approach of \citet{10.1145/566654.566623} for approximating frictional effects. 
Because SEM yields equations of motion in the standard form, it is compatible with any standard integration scheme, though in this submission we limit ourselves to
Implicit Euler.

@@

\subsection{Mass Matrix}

@@

\subsection{Error Energy}
Like all VEM-type methods, ours requires an error energy to ensure its consistency~\cite{10.1142/S021820251440003X}.
A natural choice is to adapt the error energy specified by \citet{10.1145/1073204.1073216} to penalize deviations between the pose prescribed by  $\vc{q}$ and 
that prescribed by polynomial basis.

>>>>

\begin{algorithm}[h]
    \label{alg:sem}
	\caption{Shape Matching Element Algorithm}\label{sem_algorithm}
	\begin{algorithmic}[1]
		\Procedure{PrepareSim}{$model$}
		\State{// Get initial DOF and sample boundary  \Comment{Section \ref{sec:geometric_model}}}
		\State $\vc{q} \gets \text{GetControlPoints}(model)$
		\State $\samp{\Jnurbs}, \wt \gets \text{SampleNURBS}(model)$
		\State $\vc{x^0} \gets \wt \samp{\Jnurbs} \vc{q}$
        \Statex

    	\State{// Generate material points inside the model}
		\State $\st{X} \gets \text{RaycastQuadrature}(model)$ \Comment{Section \ref{sec:quadrature}}
		\Statex

		\State{// Set of blending weights for each $\refX \in \st{X}$} 
     	\State $\st{W} \gets \text{BlendingWeights}(model, \st{X}, \alpha)$    \Comment{Section \ref{sec:weights}}
		\Statex
		\State $\XBs \gets \text{DeformationOrigins}(\st{X}, \st{w})$  \Comment{Section \ref{sec:origins}}
		\Statex

		\State{// Create projection operator}
		\State $\Pi \gets \text{ShapeMatching}(\vc{q}, \samp{\Jnurbs}, \wt, \XBs)$ \Comment{Section \ref{sec:shapematching}}
		\Statex

		\State{// Form mass matrix and error hessian, respectively}
    	\State $\MM \gets \text{MassMatrix}(\st{X}, \XBs, \st{W},\Pi)$    \Comment{Section \ref{ssec:mass_matrix}}
     	\State $\ME \gets \text{ErrorMatrix}(\vc{x^0}, \XBs, \st{W},\Pi)$  \Comment{Section \ref{ssec:error_energy}}
%     	\myworries{todo: add volume}
%     	\While {$simulating$}
%     		\State{$\Pcox \gets \Pi\mathbf{q}$}
%     		\ForAll{$i \gets 1$ to $|\st{X}|$}
%     			\State{$\mathbf{F}_i \gets \frac{\partial \mathbf{F}}{\partial c} c$}
%				\State{$\frac{\partial V}{\partial c} \gets \frac{\partial V}{\partial c} +
%				\frac{\partial \mathbf{F}}{\partial c}^T \frac{\partial \psi}{\partial \mathbf{F}}(\mathbf{F}_i)$}
%				
%				\State{$\frac{\partial^2 V}{\partial c^2} \gets \frac{\partial^2 V}{\partial c^2} +
%				 \frac{\partial \mathbf{F}}{\partial c}^T 
%				 \frac{\partial^2 \psi}{\partial \mathbf{F}^2}(\mathbf{F}_i)
%				 \frac{\partial \mathbf{F}}{\partial c}
%				 $}
%  			\EndFor
%  			\State{$K \gets \Pi^T \frac{\partial^2 V}{\partial c^2} \Pi$}
%  			\State{$f_{\text{internal}} \gets \Pi^T \frac{\partial V}{\partial c}$}
%  			\State{$f_{\text{error}} \gets \ME \vc{q}$}
%  			\State{$f_{\text{total}} \gets f_{\text{internal}} + f_{\text{error}} + f_{\text{ext}}$}
%  			\State{$\dot{\mathbf{q}} \gets \text{TimeIntegration}
%  			(f_{\text{total}},\mathbf{K}, \dot{\mathbf{q}}, \mathbf{J} , \mathbf{M}, \mathbf{M_E})$}
%  			\State{$\mathbf{q} \gets \mathbf{q} + h\dot{\mathbf{q}}$}
%     	\EndWhile
	\EndProcedure
\end{algorithmic}
\caption{Shape Matching Element Algorithm}
\end{algorithm}

%\section{Methods}
\section{Geometric Model}
\label{sec:geometric_model}

@@

\section{Physics-Based Animation}
Having defined the individual components of the SEM method, we now briefly detail how we apply it to the problem of 
physics-based animation, specifically elastodynamics. 
The advantage of having a unified kinematic description~(\refeq{kinematics}) is that we can directly apply Lagrangian Mechanics~\dave{cite Lanczos again} to arrive at 
the pertinent equations of motion:
\begin{equation*}
    \MM\ddot{\vc{q}} = \fo\left(\vc{q}\right) + \fext,
\end{equation*} where $\MM\in\real^{n\times n}$ is the mass matrix, $\vc{q}\in\real^{n}$ is the stacked vector of surface control points, $\fo\in\real^{n}$ is the 
vector of internal forces and $\fext\in\real^{n}$ are external forces such due to gravity or contact. 
We use penalty springs~\cite{10.1145/1964921.1964932} to resolve contact, and the approach of \citet{10.1145/566654.566623} for approximating frictional effects. 
Because SEM yields equations of motion in the standard form, it is compatible with any standard integration scheme, though in this submission we limit ourselves to
Implicit Euler.

@@

\subsection{Mass Matrix}
\label{ssec:mass_matrix}

@@

\subsection{Error Energy}
\label{ssec:error_energy}
Like all VEM-type methods, ours requires an error energy to ensure its consistency~\cite{10.1142/S021820251440003X}.
A natural choice is to adapt the error energy specified by \citet{10.1145/1073204.1073216} to penalize deviations between the pose prescribed by  $\vc{q}$ and 
that prescribed by polynomial basis.
